\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students

### 2023-09-19 demystify.tex
various updates

Several things about stipends.  First, the amount of stipend depends on the university, which in turns depend on various factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost). Second, a school year is (typically) 9-month in the US, so stipend is for 9 months (so divide by 9 for each month). Third, like most source of income in the US, you will have to pay tax on your stipend.  Fourth, CS department typically has higher stipend comparing to other study fields.  Finally, private universities might pay more for stipend (but they could have higher "activity" or some other hidden fee, or you will have to pay some fee for each class credits).

@@

Tenured faculty, e.g., associate and full profs., are likely older, more well-known, and have more experience in managing students.  However, they might not push you as hard (they don't have to, they already got tenure). They might also expect you to figure things out yourself, i.e., so you need to be very independent.  Some tenured faculty are also no longer active in research and more involved with administrative duties or with their startup companies (this means they will likely not take new students).

You can learn about faculty and their level of research activity through the faculty's website or even from CSRankings.
Ultimately, choose one that fits you the most by communicating with them, meeting them, asking them questions, and even talking to their students.

@@

\begin{enumerate}
    \item \textbf{Cheating}: Cheating, e.g., exams, coding, is a big no-no in the US.  If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time at GMU).   This is very different from many international countries where cheating is common and often tolerated.  Faculty is extremely good at detecting cheating (we have been dealing with these situations so many times over so many years), and \emph{will} report cheating cases.  In short, whatever you do, don't cheat---not worth it.

    \item \textbf{Illegal Software}: Using illegal/cracked software is very common in many countries. But do not use/install them on university computer (e.g., computers given by your department or advisor).  It is unlikely that the university will track you down, but it is the \textbf{software company} that will.  They have very sophisticated tools to detect illegal software and will sue your university/department.  Imagine your department or advisor being sued for half a million dollars, and it is you who caused it.   If you need software, ask your advisor or the department.
\end{enumerate}

>>>>

Several things about stipends.  First, the amount of stipend depends on the university, which in turns depend on various factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost). Second, a school year is (typically) 9-month in the US, so stipend is for 9 months (so divide by 9 for each month). Third, like most source of income in the US, you will have to pay tax on your stipend.  Fourth, CS department typically has higher stipend comparing to other study fields.  Finally, private universities might pay more for stipend (but they could have higher "activity" or some other hidden fee, or you may be required to pay fees for each credit hour).

@@

Tenured faculty, e.g., associate and full profs., are likely older, more well-known, and have more experience in managing students.  However, they might not push you as hard (they don't have to, they already got tenure). They might also expect you to figure things out yourself, i.e., so you need to be very independent.  Some tenured faculty are also no longer active in research and more involved with administrative responsibilities or with their startup companies (this means they will likely not take new students).

You can learn about faculty and their level of research activity through the faculty's website or even from CSRankings.
Ultimately, choose one that fits you the most by communicating with them, meeting them, asking them questions, and even talking to their current students.

@@

\paragraph{Diversity} US universities prioritize diversity and inclusion. It's crucial for students to respect and appreciate varied opinions, backgrounds, and cultures. Unlike some countries where certain voices are marginalized, in the US, all perspectives are valued equally (especially at universities). Racism or discrimination can have serious consequences, including academic and disciplinary actions.

\paragraph{Academic Integrity} Plagiarism and cheating, e.g., exams, coding, is a big no-no in the US.  If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time at GMU).   This is very different from many international countries where cheating is common and often tolerated.  Faculty is extremely good at detecting cheating (we have been dealing with these situations so many times over so many years), and \emph{will} report cheating cases.  In short, whatever you do, don't cheat---not worth it.

Here is the typically steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university;  (ii)  OAI will completely take over and investigate the case; and finally (iii) OAI make the final decision.  It is important to note that after receiving the report from your prof., OAI completely takes over (so begging your professor will not help because they simply are no longer involved).

\paragraph{Illegal Software} Using illegal/cracked software is very common in many countries. But do not install or use them on university computer (e.g., computers given by your department or advisor).  It is unlikely that the university will track you down, but it is the \textbf{software company} that will.  They have very sophisticated tools to detect illegal software and will sue your university/department.  Imagine your department or advisor being sued for half a million dollars, and it is you who caused it.   If you need software, ask your advisor or the department.
